Uncertainty propagation in deep neural networks for sparse coding

### 2018-03-10 paper/BayesianLISTA.tex
image size reduced and captions amended

\section{\uppercase{Experiments}}
\label{sec:experiments}
The proposed Bayesian \textsc{lista} is evaluated in the context of the sparse coding problem with underdefined dictionary, where the number of measurements $K$ is much smaller than the dimensionality of the vector $\boldsymbol\beta$.

@@

\begin{figure}[t]
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=\columnwidth]{graphics/synthetic_number_of_layers/nmse_validation}}
\newline
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=\columnwidth]{graphics/synthetic_number_of_layers/f_measure_validation}}
\caption{Number of layers on synthetic data}
\label{fig:number_of_layers_synthetic}
\end{figure}

\begin{figure}[t]
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=\columnwidth]{graphics/synthetic_undersampling/nmse_validation}}
\newline
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=\columnwidth]{graphics/synthetic_undersampling/f_measure_validation}}
\caption{Size of observations on synthetic data}
\label{fig:unsersampling_synthetic}
\end{figure}

@@

\begin{figure}[t]
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=\columnwidth]{graphics/mnist/100_normalised_nmse_valid}}
\newline
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=\columnwidth]{graphics/mnist/100_normalised_f_measure_valid}}
\caption{\textsc{mnist} with dictionary $100$}
\label{fig:mnist_100}
\end{figure}

\begin{figure}[t]
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=\columnwidth]{graphics/mnist/250_normalised_nmse_valid}}
\newline
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=\columnwidth]{graphics/mnist/250_normalised_f_measure_valid}}
\caption{\textsc{mnist} with dictionary $250$}
\label{fig:mnist_250}
\end{figure}

%\begin{figure}[t]
%\includegraphics[width=\columnwidth]{loss}
%\caption{Validation \textsc{nmse}}
%\label{fig:validation}
%\end{figure}
\begin{figure}[t]
\subfloat[$\beta$ posterior mean]{\includegraphics[width=0.33\columnwidth]{posterior_mean}}~
\subfloat[$\beta$ posterior std]{\includegraphics[width=0.33\columnwidth]{posterior_std}}~
\subfloat[$\beta$ posterior spike indicator]{\includegraphics[width=0.33\columnwidth]{posterior_spike_indicator}}
\caption{Posterior for the digit 7.}
\label{fig:posterior_distribution}
\end{figure}
\begin{figure}[t]
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_0}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_1}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_2}}
\caption{Samples from the posterior for the digit 7.}
\label{fig:posterior_samples}
\end{figure}

\subsection{\uppercase{Active learning on mnist}}
To demonstrate the potential scenario that can benefit from uncertainty estimates of the Bayesian \textsc{lista} we consider the active learning example \citep{settles.tr09}. Active learning area researches ways to select new training subsets to reduce total number of required supervision. One of the approaches for active learning is uncertainty sampling when the data with the least certain predictions is used for training. The uncertainty can be measured with entropy or variance. In our example no closed form for entropy of spike and slab exist, therefore we use variance from lemma \ref{thm:moments_spsl} as a measure of uncertainty.

@@

The performance on the actively updated train data is compared with the performance on the randomly updated train data, when at every iteration the random element is moved from pool set to train set. Both algorithms of new data sampling were run with $20$ random seeds and their performance was averaged. Figure \ref{fig:active_learning_mnist} demonstrates their performance.The uncertainty sampling has some advantage compared to random sampling, this means that the posterior variance is adequately estimated.
\begin{figure}[t]
\subfloat[\textsc{nmse}]{\includegraphics[width=\columnwidth]{graphics/active_mnist/nmse_validation}} \\
\subfloat[F measure]{\includegraphics[width=\columnwidth]{graphics/active_mnist/f_measure_validation}}
\caption{Active learning on \textsc{mnist}.}
\label{fig:active_learning_mnist}
\end{figure}

>>>>

\section{\uppercase{Experiments}}
\label{sec:experiments}
The proposed Bayesian \textsc{lista} is evaluated in the context of the sparse coding problem with an undercomplete dictionary, where the number of measurements $K$ is much smaller than the dimensionality of the vector $\boldsymbol\beta$.

@@

\begin{figure}[t]
\centering
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/nmse_validation}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/f_measure_validation}}
\caption{Predictive accuracy on different number of layers (for neural networks) or iterations (for baselines) on the synthetic data}
\label{fig:number_of_layers_synthetic}
\end{figure}

\begin{figure}[t]
\centering
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/nmse_validation}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/f_measure_validation}}
\caption{Predictive performance on different size of observations on the synthetic data}
\label{fig:unsersampling_synthetic}
\end{figure}

@@

\begin{figure}[t]
\centering
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_nmse_valid}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/100_normalised_f_measure_valid}}
\caption{Predictive performance on number of iterations on the \textsc{mnist} data with dictionary size $K = 100$}
\label{fig:mnist_100}
\end{figure}

\begin{figure}[t]
\centering
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_nmse_train}}
%~
\subfloat[\textsc{nmse} on validation]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_nmse_valid}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_f_measure_train}}
%~
\subfloat[F measure on validation]{\includegraphics[width=0.5\columnwidth]{graphics/mnist/250_normalised_f_measure_valid}}
\caption{Predictive performance on number of iterations on the \textsc{mnist} data with dictionary size $K = 250$}
\label{fig:mnist_250}
\end{figure}

%\begin{figure}[t]
%\includegraphics[width=\columnwidth]{loss}
%\caption{Validation \textsc{nmse}}
%\label{fig:validation}
%\end{figure}
\begin{figure}[t]
\centering
\subfloat[$\beta$ posterior mean]{\includegraphics[width=0.33\columnwidth]{posterior_mean}}~
\subfloat[$\beta$ posterior std]{\includegraphics[width=0.33\columnwidth]{posterior_std}}~
\subfloat[$\beta$ posterior spike indicator]{\includegraphics[width=0.33\columnwidth]{posterior_spike_indicator}}
\caption{Posterior parameters for an image of digit 7.}
\label{fig:posterior_distribution}
\end{figure}
\begin{figure}[t]
\centering
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_0}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_1}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_2}}
\caption{Samples from the posterior for an image of digit 7.}
\label{fig:posterior_samples}
\end{figure}

\subsection{\uppercase{Active learning}}
To demonstrate the potential scenario that can benefit from uncertainty estimates of the Bayesian \textsc{lista} we consider the active learning example \citep{settles.tr09}. Active learning area researches ways to select new training subsets to reduce total number of required supervision. One of the approaches for active learning is uncertainty sampling when the data with the least certain predictions is used for training. The uncertainty can be measured with entropy or variance. In our example no closed form for entropy of spike and slab exist, therefore we use variance from lemma \ref{thm:moments_spsl} as a measure of uncertainty.

@@

The performance on the actively updated train data is compared with the performance on the randomly updated train data, when at every iteration the random element is moved from pool set to train set. Both algorithms of new data sampling were run with $20$ random seeds and their performance was averaged. Figure \ref{fig:active_learning_mnist} demonstrates their performance.The uncertainty sampling has some advantage compared to random sampling, this means that the posterior variance is adequately estimated.
\begin{figure}[t]
\centering
\subfloat[\textsc{nmse}]{\includegraphics[width=0.5\columnwidth]{graphics/active_mnist/nmse_validation}}%
\subfloat[F measure]{\includegraphics[width=0.5\columnwidth]{graphics/active_mnist/f_measure_validation}}
\caption{Performance in the active learning experiment on the \textsc{mnist} data. }
\label{fig:active_learning_mnist}
\end{figure}
